Basic Representation Types\\ \Large Natural Language Processing Toolkit: Technical Report

### 2001-05-12 technical/reptypes/reptypes.tex
Added lots of content.. a paragraph about locations, wrote the section on words, most of the section on collectino of words..
svn/trunk@30

@@

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath}

@@

  Section \ref{sec:concepts} describes the basic concepts which this
  technical report attempts to address.  Sections \ref{sec:locations}
  through \ref{sec:trees} describe the basic types used to encapsulate
  these concepts.  Section \ref{sec:summary} provides a summary of the
  basic types, and discusses how they interact.

@@

  The location of an element of text is represented using a span-based
  data structure.  A \emph{span} consists of a \emph{start index} and
  an \emph{end index}, both of which are integers.  A span with start
  index $s$ and end index $e$ is written ``$[s,e)$,'' and specifies
  the location of the text beginning at $s$, and including everything
  up to (but not including) the text at $e$.  It is always true that
  $s<e$.\footnote{Should this be changed to $s\leq e$?}

@@

    \item A \emph{start index} (accessed via the \code{start()} member
    function).  \\Type: \code{int}

    \item An \emph{end index} (accessed via the \code{end()} member
    function).  \\Type: \code{int}

    \item An optional \emph{unit} (accessed via the \code{unit()}
    member function).  \\Type: \code{string} \emph{(normalized to
    lower case)}

    \item An optional \emph{source} (accessed via the \code{source()}
    member function).  \\Type: \emph{(any)}

@@

  \noindent
  Locations are immutable objects.

  Locations can be compared for equality and ordering.  If two
  locations with different units or different sources are compared,
  then an exception will be raised.  When location comparisons do not
  raise exceptions, they are defined as follows:
  \begin{align*}
     [s_1,e_1) = [s_2,e_2) &\iff s_1=s_2 \quad \land \quad e_1=e_2 \\
     [s_1,e_1) < [s_2,e_2) &\iff e_1 < s_2 \\
     [s_1,e_1) > [s_2,e_2) &\iff s_1 > e_2 \\
     [s_1,e_1) \leq [s_2,e_2) &\iff [s_1,e_1) < [s_2,e_2) \quad \lor \quad
                             [s_1,e_1) = [s_2,e_2) \\
     [s_1,e_1) \geq [s_2,e_2) &\iff [s_1,e_1) > [s_2,e_2) \quad \lor \quad
                             [s_1,e_1) = [s_2,e_2) 
  \end{align*}

  \noindent
  Note that if two locations $l_1$ and $l_2$ overlap, then it will
  never be the case that $l_1<l_2$ or $l_1>l_2$.

  \subsection{Explanation and Comments}

@@

  We decided to restrict location indices to integers because it
  simplifies the interface without signifigantly reducing its power.
  Other alternatives we considered were floating point numbers and
  arbitrary comperable objects.  Floating point numbers were rejected
  because rounding errors can invalidate tests for equality and
  ordering.  If a floating-point index seems appropriate, it is
  recommended that users either multiply the floating point numbers by
  a fixed constant and round them to integers, or create a
  bidirectional mapping between floating-point indices and integer
  indices.  The use of arbitrary comperable objects as indices was
  rejected because it signifigantly complicates the location
  interfaces, without adding signifigant power to the system.  Also,
  arbitrary comperable objects may incur some of the difficulties that
  would be encountered with floating point numbers.

@@

@@

@@

@@

\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The abstract concepts which

This document catalogs some of the different conceptual entities which
the tooklit should deal with.  These conceptual entities sometimes
draw more distinctions than is convenient when actually working with
objects.

An open question is how strict the toolkit should be about enforcing
the distinctions between conceptually distinct types of entities.  On
one hand, strictly enforcing these distinctions is conceptually
``simpler,'' and may promote clearer thinking.  On the other hand, it
is often cumbersome to make these distinctions -- in many
circumstances, they don't matter, so they get dropped in practice.
For example, most people will not distinguish ``tokens'' from ``token
types'' unless they need to.  In a sense, this question boils down to
whether it will be easier to learn/use a set of rules/conventions that
are simpler, but somewhat less intuitive; or more complex, but more
intuitive.  I think that one danger is that the more complex rules
will only be more intuitive some of the time.  And it may be dangerous
to be inconsistantly intuitive..

Another open question is whether the current names should be kept.  In
particular, names of conceptual entities should be both \emph{correct}
and \emph{easy to use/remember}.

Also, it may be possible that some of these conceptual entities could
be merged or done away with altogther.  Are all of these entities
serving useful purposes?

\section{Token Stuff}

{\it (Currently, many of these definitions
are a bit hand-wavey.  That partially stems from the fact that they
are not really as well-defined as we would like (esp the more
primitive ones).  But I'll try to clean them up eventually.)}

  \item \textbf{Lexeme:} An identifier for a word-like thing.
  Normally, this would be a string.  But if you wanted to process text
  more efficiently, it could be an integer.  It might be stemmed, not
  stemmed, case-normalized, not case-normalized, etc.

  \item \textbf{Token Type:} A word-like unit of text.  Includes a
  lexeme, and may also include other info, like part of speech tag,
  etc.

  \item \textbf{Token:} An occurance of a token type

  \item \textbf{Location:} The location of something in text.  There
  are actually a number of ways of representing locations: as indexes,
  as spans, etc.  Also, indexes can refer to word number, character
  number, etc.

\section{Syntax Tree Stuff}

  \item \textbf{SyntaxNode:} The annotation that appears at a node
  in a syntax tree.  E.g., this might hold a phrase type (``NP'' or
  ``V'').

  \item \textbf{SyntaxTree:} An occurance of a tree or subtree.
  Consists of a SyntaxNode and an ordered list of zero or more
  children.  Children must be SyntaxTrees or Tokens.

  \item \textbf{SyntaxTreeType:} The ``type'' of a syntax tree (as
  opposed to an occurance of a syntax tree).  Consists of a SyntaxNode
  and an ordered list of zero or more children.  Children must be
  SyntaxTreeTypes or TokenTypes.

\section{How much to enforce?}

So basically, the question of deciding how strictly to enforce these
distinctions comes down to two questions:

\begin{enumerate}
  \item Should any entities/distinctions be eliminated?
  \item Should it be possible to use entities in environments that
        are not \emph{strictly speaking} correct?
\end{enumerate}

I'll leave the first question for later, and concentrate on the
second.  There are a number of ways that objects could try to be more
friendly when you use them in ways that that are not strictly speaking
correct:

\begin{itemize}
  \item Equality tests: What happens when you compare a string to a
  token?  A string to a TokenType?  A Token to a TokenType?  etc.?  A
  related question is how to handle comparisons within one class
  hierarchy.  E.g., equality tests between a tagged token and a simple
  token.  The strict view would say that any of these comparisons
  should raise an exception.  However, if we decide not to be strict,
  there are a number of places where these equality tests could be
  well-defined.  Examples that come to mind are:
  \begin{itemize}
    \item equality between a string and a SimpleTokenType
    \item equality between a string and a Token (ignore location)
    \item equality between an integer and a Location
    \item equality between a list and a SyntaxTree(Type)
  \end{itemize}

  \item Arguments to methods: Should methods attempt to handle
  arguments that are not quite the expected type?  E.g., when Steven
  passed strings to a FreqDist, when it was really (sort of) expecting
  TokenTypes.

  \item Arguments to constructors: Should constructors automatically
  ``wrap'' objects if they're not the right type?  E.g., should
  Token(``cat'') automatically create a SimpleTokenType containing
  ``cat''?
\end{itemize}

>>>>

% Note: be careful to say ``x is a kind of y'' instead of ``x is a type 
% of y,'' since we're giving ``type'' special meaning.

@@

\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{amsmath}

@@

  Section \ref{sec:concepts} describes the basic concepts which this
  technical report attempts to address.  Sections \ref{sec:locations}
  through \ref{sec:trees} describe the basic types that were chosen to
  encapsulate these concepts, and the rational behind choosing those
  types.  Section \ref{sec:summary} provides a summary of the basic
  types, and discusses how they interact.

@@

  The location of an element of text is represented using a span-based
  data structure.  A \emph{span} consists of a \emph{start index} and
  an \emph{end index}, both of which are non-negative integers.  A
  span with start index $s$ and end index $e$ is written ``$[s,e)$,''
  and specifies the location of the text beginning at $s$, and
  including everything up to (but not including) the text at $e$.  It
  is always true that $s<e$.\footnote{Should this be changed to $s\leq
  e$?  If $s=e$, then the span would be empty -- does it make sense to
  have empty locations?}

@@

    \item A \emph{start index} (accessed via the \code{start()} member
    function).  \\
    Type: \code{int}

    \item An \emph{end index} (accessed via the \code{end()} member
    function).  \\
    Type: \code{int}

    \item An optional \emph{unit} (accessed via the \code{unit()}
    member function).  \\
    Type: \code{string} \emph{(normalized to lower case)}

    \item An optional \emph{source} (accessed via the \code{source()}
    member function).  \\
    Type: \emph{(any)}

@@

  \noindent
  \code{Location}s are immutable objects.  \code{Location}s can be
  compared for equality and ordering.  If two \code{Location}s with
  different units or different sources are compared, then an exception
  will be raised.  When \code{Location} comparisons do not raise
  exceptions, they are defined as follows:
  \begin{align*}
     [s_1,e_1) = [s_2,e_2) &\iff s_1=s_2 \quad \land \quad e_1=e_2 \\
     [s_1,e_1) < [s_2,e_2) &\iff e_1 < s_2 \\
     [s_1,e_1) > [s_2,e_2) &\iff s_1 > e_2 \\
     [s_1,e_1) \leq [s_2,e_2) &\iff [s_1,e_1) < [s_2,e_2) \quad \lor \quad
                             [s_1,e_1) = [s_2,e_2) \\
     [s_1,e_1) \geq [s_2,e_2) &\iff [s_1,e_1) > [s_2,e_2) \quad \lor \quad
                             [s_1,e_1) = [s_2,e_2) 
  \end{align*}

  \noindent
  Note that if two \code{Location}s $l_1$ and $l_2$ overlap, then it
  will never be the case that $l_1<l_2$ or $l_1>l_2$.

  \subsection{Discussion}

@@

  We decided to restrict location indices to non-negative integers
  because it simplifies the interface without signifigantly reducing
  its power.  Other alternatives we considered were floating point
  numbers and arbitrary comperable objects.  Floating point numbers
  were rejected because rounding errors can invalidate tests for
  equality and ordering.  If a floating-point index seems appropriate,
  it is recommended that users either multiply the floating point
  numbers by a fixed constant and round them to integers, or create a
  bidirectional mapping between floating-point indices and integer
  indices.  The use of arbitrary comperable objects as indices was
  rejected because it signifigantly complicates the location
  interfaces, without adding signifigant power to the system.  Also,
  arbitrary comperable objects may incur some of the difficulties that
  would be encountered with floating point numbers.

  A span $[s,e)$ includes the text at $s$ but not the text at $e$ in
  order to be consistant with Python slice notation.  Also, all
  indices are required to be non-negative for consistancy with Python
  slice notation, where negative indices have special meaning.

@@

  The term ``word'' is used to refer to a wide variety of different
  abstract concepts, as can be seen by looking through the list of
  concepts from Section \ref{sec:concepts}.  One important distinction
  that is easy to overlook is the difference between a word and an
  individual occurance of a word.  For example, the sentence ``my dog
  likes his dog'' contains four words but five occurances of words
  (since the word ``dog'' appears twice).  In order to make this
  distinction clear, we will use the words \emph{type} and
  \emph{token} to refer to words and to occurances of words,
  respectively.  For example, in the sentence ``my dog likes his
  dog,'' we can say that the second and fifth words are different
  tokens, but that they have the same types.

  Individual tokens extracted from the same text can always be
  distinguished based upon their locations.  This fact provides us
  with a convenient mechanism for encoding tokens: a token is simply a
  pair consisting of a type and a location.  Two tokens are only equal
  if both their location and thier type are equal.

  The notions of token and type are actually more general than the
  notions of word and word occurance.  In the context of the toolkit,
  types can refer to any object derived from text, and tokens can
  refer to any instances of objects derived from text.  For example,
  syntax trees are actually special kinds of tokens (see Section
  \ref{sec:trees}).

  The toolkit does not define a specific encoding for individual kinds
  of tokens or token types, like simple words, tagged words, or
  stemmed words.  Instead, it defines the general encoding for tokens,
  and allows students to use almost any object as a type.  For
  example, \code{string}s could be used to encode the types of simple
  words, stemmed words, and case-normalized words; and pairs of
  strings could be used to encode tagged word types.

  However, the toolkit does provide a number of simple classes that
  are intended to be used as types.  Currently, the only class
  provided by the toolkit is the \code{TaggedType} class, which can be
  used to encode the type of a tagged token.  A \code{TaggedType}
  simply consists of a base type and a tag, both of which are
  \code{string}s.

  \subsubsection{Types}
      A Type can be any immutable object that properly implements
      equality comparison.

  \subsubsection{Tokens}
      Tokens are implemented using the \code{Token} class.
      \code{Token}s consist of:

        \item A \emph{location} (accessed via the \code{loc()} member
        function).  \\
        Type: \code{Location}

        \item A \emph{type} (accessed via the \code{type()} member
        function).  \\
        Type: \emph{(any)}

      \noindent
      \code{Token}s are immutable objects.

  \subsubsection{TaggedTypes}
      \vspace{3mm}\noindent
      The types of tagged tokens can be encoded using the
      \code{TaggedType} class.  A \code{TaggedType} consists of:

        \item A \emph{base type} (accessed via the \code{base()}
        member function).  \\
        Type: \code{string}

        \item A \emph{tag} (accessed via the \code{tag()} member
        function).  \\
        Type: \code{string}

      \noindent
      \code{TaggedType}s are immutable objects.  Two tagged types are
      equal if their bases are equal and their tags are equal.  Note
      that equality of strings is defined to be case-sensitive.

  \subsection{Discussion}
  % - why not a class/interface for type?
  % - why location+type?
  % - why not classes to distinguish stemmed/case-normalized/etc?

  We considered implementing an interface or class hierarchy for
  types.  Using an interface, rather than allowing arbitrary Python
  objects, has a number of benefits:

  \begin{description}

    \item[Exception Semantics] If an interface is used to define what
    type objects must do, then types can exhibit more precise
    exception semantics.  In particular, it would be possible to
    ensure that comparing two type objects with different classes
    would always raise an exception.  In constrast, it is impossible
    to make such guarantees when comparing arbitrary Python objects.

    \item[Type-kind Distinctions] If each kind of type (e.g., stemmed
    types, case-normalized types, tagged types, etc.) is implemented
    with a separate class, then it is easier to maintain distinctions
    between different kinds of types.  But if built-in Python objects
    such as strings and tuples are used to represent types, then the
    distinctions between different kinds of type will become blurred.
    For example, since strings are likely to be used for simple words,
    stemmed words, and case-normalized words, it will be impossible to
    determine what kind of type a string contains by inspecting the
    string object.

  \end{description}

  However, requiring that all types implement an interface often
  introduces an extra layer of indirection, and complicates the system
  used to represent words.  Since words are fundamental to the
  toolkit, we decided that they should be kept as simple as is
  practical.  Some of the exception semantics can be regained at a
  different level.  For example, even though testing two types for
  equality may not raise an exception if they have different Python
  types, it would be possible to implement the \code{type\_eq()}
  method of tokens to perform this check.\footnote{The
  \code{type\_eq()} method is also included because it can be more
  efficient than extracting the tokens' types and comparing them,
  especially for some special kinds of tokens like trees (see Section
  \ref{sec:trees}).}  Also, although type-kind distinctions cannot be
  derived directly from the type object, these distinctions are
  usually implicit in the history of the object, and it should be rare
  to have a type object without knowing what kind of type it contains.
  Therefore, we decided that the benefits derived from maintaining a
  class hierarchy for types did not justify the increase in complexity
  for the toolkit's set of basic type.

@@

  The toolkit does not implement any special encodings for collections
  of words.  Instead, it relies on Python's built-in types.  However,
  there are a number of standard ways to use Python's built-in types
  to encode collections of words:

  \begin{description}

    \item[Strings] Unprocessed text is usually represented as a single
    string.  For example, when data is read from a file, it is
    typically read into a string.  Tokenizers can be used to convert
    strings into lists of tokens.

    \item[List of Tokens] A piece of text can be represented as an
    ordered list of the tokens that it contains.

        \item A list of tokens $l$ is said to be ``properly ordered''
        if, for every $i<j$, $l[i].loc()<l[j].loc()$.

        \item A list of tokens $l$ is said to be ``exhaustive'' if:

        \begin{itemize}
          \item for every $i$, $l[i].end() = l[i+1].start()$; 
          \item $l[0].start()$ is the start of the text.
          \item $l[-1].end()$ is the end of the text.
        \end{itemize}

    \item[List of Types] When the locations of individual tokens is
    unimportant, a text can be represented as an ordered list of its
    token's types.

    \item[Set of Tokens] When the order of the tokens in a text is
    unimportant, the text can be represented as a set of tokens.
    \footnote{Python does not (yet) provide a built-in set type;
    however, the toolkit defines a generic set class.}  Since each
    token's location serves to distinguish it from all other tokens,
    the set will not contain any duplicate elements.

    \item[Set of Types] When the order of tokens in a text and the
    presence of multiple tokens with the same type are unimportant,
    the text can be represented as a set of types.

    \item[Dictionary Mappings] Since both tokens and types are
    immutable objects, they can be used as keys in dictionaries.  A
    typical use of a dictionary would be to encode a multiset of types
    as a dictionary mapping types to counts.

  \end{description}

%  \subsection{Implementation}
%  There is no real implementation, is there??

  \subsection{Discussion}

  % Why not explicit encodings?  Advantage of explicit encodings:
  % check that everything in a list/set/whatever has the same pytype.

@@

  \subsection{Discussion}

@@

%% Meta-section
\newpage
\section{Meta-Questions}
\begin{itemize}
  \item Are the terms used in this document correct?  Could better
  terms be chosen?  Especially: properly ordered, exhaustive,
  type-kind.
  \item Are the interfaces presented here reasonable?
  \item Do any decisions need more justification?
\end{itemize}

\end{document}
